Distributional Trees and Forests

### 2017-11-15 pkg/disttree/inst/draft/draft.Rnw
edited plots, small corrections
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1709 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 46dd1abe8808febd4c056e4e8fe58ed62a22c2b4
changed setup for predictions, now learn data set consists of 24 years and predictions are made for 4 years
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1714 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: a45939a53927c461a631b2b5fabc3a24da2ef2c5

After explaining the novel methodology in Section~\ref{Methodology} of this paper and presenting results from a simulation study in Section~\ref{Simulation} the focus will then be put on an application in the field of probabilistic forecasting in Section~\ref{ProbabilisticForecasting}. The new method is applied on precipitation data in a mountainous region with a large number of numerical weather prediction quantities and the results are compared to those of other methods commonly used in this research area. In a first setting only one observation station is taken into consideration. The models are trained on 25 years and prbabilistic predictions of the consecutive 3 years are validated. A brief outlook on the results of this setting is given by Figure~\ref{axams_onestation}. For the 24th of July of 2010, 2011 and 2012 the density functions of the distributions estimated by a forest model are plotted together with the corresponding observations. This illustrates that well fitting predicted expected values are provided by this model as they are close to the observed values.
Beside the expected values the predictions provided by distributional models offer a wide range of information as they specify the whole distribution and therefore enable further methods of inference and analysis.

@@

  save(res, file = "rain_Axams_pred.rda")
}

@@

}
@

\begin{figure}[t!]
\begin{center}
<<rain_Axams_pred, fig=TRUE, echo=FALSE, width=6.5, height=4.5>>=
plot(x = x, y = y1, type = "l", col = "red", 
     main = paste0("July ", pday, dayending), ylab = "Density", 
     #xlab = expression(Total~precipitation~group("[", list(mm^(1/1.6), plain(per 24h)),"]")),
     xlab = expression(Total~precipitation~group("[", mm^(1/1.6)~per~day,"]")),
     #{\frac{1}{1.6}}$,
     ylim = c(0,max(y1, y2, y3, y4, pm1, pm2, pm3, pm4) + 0.01),
     xlim = c(-1.5,8))

lines(x = x, y = y2, type = "l", col = "blue")
lines(x = x, y = y3, type = "l", col = "darkgreen")
lines(x = x, y = y4, type = "l", col = "purple")
legend('topright', c("predicted distribution", "point mass at censoring point", "observation"),
       bty = "n", col = "black", lty = c(1, NA, NA), pch = c(NA, 19, 4), cex = 1)

@@

# add labels
text(x = -0.8, y = lh1, labels = "2009", col = "red", cex = 0.8)
text(x = -0.8, y = lh2, labels = "2010", col = "blue", cex = 0.8)
text(x = -0.8, y = lh3, labels = "2011", col = "darkgreen", cex = 0.8)
text(x = -0.8, y = lh4, labels = "2012", col = "purple", cex = 0.8)
@
\caption{\label{axams_onestation}Predictions at station Axams for one day in July of 2010, 2011 and 2012. Observations and predictions are left censored at 0. The corresponding point mass is shown at the censoring point (0).}
\end{center}
\end{figure}

@@

\subsection{Motivation (Review/Problems)}
As illustrated in Figure~\ref{Devel_parmod} a lot of development has been made in the field of parametric modeling and a wide range of tools can be applied in research areas such as probabilistic forecasting. But as mentioned before, further features such as the ability to capture non-additive effects or interactions are desirable in many applications which demands for more flexible models. Since tree structured models provide these required features a combination of them with parametric models can offer a framework that is able to deal with a wide range of different modeling situations. In this section we are going to introduce the novel methodology which embeds parametric modeling into the idea of tree structured models step by step. We start with a simple distributional fit in Subsection~\ref{distfit}, then build a distributional tree in Subsection~\ref{disttree} and finally a distributional forest in Subsection~\ref{distforest}.

@@

The distribution of the response variable is a left censored normal distribution with censoring point 0. The corresponding distribution parameters are defined by a parameter function $f$ depending on the covariates and an additional factor $\kappa$, also known as the smoothness constraint, which regulates the steepness of certain sections of the parameter function. For low values of $\kappa$ the resulting parameters change rather slowly over varying values of the covariates while high values of $\kappa$ lead to abrupt changes. \\ 
The location parameter $\mu$ depends on $X_1, X_4$ and $\kappa$ and the scale parameter $\sigma$ depends on $\mu, X_1$ and $\kappa$. The other covariates $X_2, X_3, X_5$ and $X_6$ are used as noise variables.\\
The latent variable $Y^{\ast}$ is generated based on the following setting.
\begin{center}
$$ 
Y^{\ast} \sim \mathcal{N} (f(X_1, X_4, \kappa))
$$
with
$$
f(X_1, X_4, \kappa) = \left(\mu(X_1, X_4, \kappa) , \sigma(\mu,X_1,\kappa)\right)
$$ 
and
\begin{align*}
\mu(X_1, X_4, \kappa) = 0.2 
+ \begin{cases}
    7 \cdot \exp{(-(3 \cdot X_1-1.2)^{(2\cdot \kappa)})},
    & \text{if } X_1 < 0.4 \text{ and } X_4 = 0\\
    7 \cdot \exp{(-(3 \cdot X_1-1.2)^{(2\cdot \kappa)})} \cdot \frac{1}{2} + \frac{7}{2},       
    & \text{if } X_1 \geq 0.4 \text{ and } X_4 = 0\\
    10 \cdot \exp{(-(3 \cdot X_1-1.2)^{(2\cdot \kappa)})},
    & \text{if } X_1 < 0.4 \text{ and } X_4 = 1\\
    10 \cdot \exp{(-(3 \cdot X_1-1.2)^{(2\cdot \kappa)})} \cdot \frac{1}{2} + \frac{10}{2},       
    & \text{if } X_1 \geq 0.4 \text{ and } X_4 = 1
\end{cases}
\end{align*}

$$
\sigma(\mu, X_1,\kappa) = 0.5 + \frac{\mu}{5} + 2 \cdot \mathbf{1}(X_1 < -0.7)
$$
%where $\Lambda$ is the cumulative distribution function of the logistic distribution
%$$
%\Lambda(\omega) = \frac{1}{1 + \exp(-\omega)}
%$$
The response variable $Y$ is defined as 
$$
Y = 
\begin{cases}
    Y^{\ast},& \text{if } Y^{\ast} \geq 0\\
    0,       & \text{else}
\end{cases}
$$
\end{center}

To illustrate the parameter functions and the impact of $\kappa$ together with the results of the considered methods they are first applied on one generated data set only where $\kappa$ is set to 1 and then to a second data set where $\kappa$ is set to 10. In Figure~\ref{plot_sim_oneset_mu} the location parameter function $\mu(X_1, X_4, \kappa)$ is plotted along the covariate $X_1$ together with the estimated values of the different models for fixed covariates $x_4 = 1$ and $(x_2, x_3, x_5, x_6) = (0,0,0,0)$. \\

@@

  #define parameter function
  kappa <- 1
  fun1 <- function(x){
    mu <- 0.2 + 
      (7 * (exp(-(3*x[,1]-1.2)^(2*kappa)))) * as.numeric(x[,1] < 0.4) + 
      (7 * (exp(-(3*x[,1]-1.2)^(2*kappa))) * (1/2) + 7/2) * as.numeric(x[,1]>= 0.4) +
      (3 * (exp(-(3*x[,1]-1.2)^(2*kappa)))) * x[,4]
    sigma <- 0.5 + 
      mu/5 + 
      #(x[,1]>0.4)  #*x[,4]#+
      2*(x[,1] < -0.4) 
      #plogis((kappa^(1.8)) * 4 * (x[,2] + 0.7)) #* x[,4]
      #4*(1-plogis((kappa^(1.8)) * 2 * (x[,1] + 0.5))) * x[,4]
      #5 * abs(x[,1]) * x[,4]
    par <- cbind(mu, sigma)
    return(par)
  }

@@

  oneset1 <- sim_oneset_vary_x1(parfun = fun1, nobs = 500, seedconst = 74, ntree = 100,
                                 formula = y~x1+x2+x3+x4+x5+x6, 
                                 tree_minsplit = 20, tree_minbucket = 7, tree_mincrit = 0.95, 
                                 forest_minsplit = 20, forest_minbucket = 7, forest_mincrit = 0, 
                                 forest_mtry = 3, type.tree = "ctree", 
                                 censNO = TRUE, 
                                 pred_fix_x2 = 0.9,
                                 pred_fix_x4 = 1,
                                 nrunif = 3,
                                 nrcovform = 6)

@@

  #define parameter function
  kappa <- 10
  fun10 <- function(x){
    mu <- 0.2 + 
      (7 * (exp(-(3*x[,1]-1.2)^(2*kappa)))) * as.numeric(x[,1] < 0.4) + 
      (7 * (exp(-(3*x[,1]-1.2)^(2*kappa))) * (1/2) + 7/2) * as.numeric(x[,1]>= 0.4) +
      (3 * (exp(-(3*x[,1]-1.2)^(2*kappa)))) * x[,4]
    sigma <- 0.5 + 
      mu/5 + 
      #(x[,1]>0.4)  #*x[,4]#+
      2*(x[,1] < -0.4) 
      #plogis((kappa^(1.8)) * 4 * (x[,2] + 0.7)) #* x[,4]
      #4*(1-plogis((kappa^(1.8)) * 2 * (x[,1] + 0.5))) * x[,4]
      #5 * abs(x[,1]) * x[,4]
    par <- cbind(mu, sigma)
    return(par)
  }

@@

  oneset10 <- sim_oneset_vary_x1(parfun = fun10, nobs = 500, seedconst = 74, ntree = 100,
                                 formula = y~x1+x2+x3+x4+x5+x6, 
                                 tree_minsplit = 20, tree_minbucket = 7, tree_mincrit = 0.95, 
                                 forest_minsplit = 20, forest_minbucket = 7, forest_mincrit = 0, 
                                 forest_mtry = 3, type.tree = "ctree", 
                                 censNO = TRUE, 
                                 pred_fix_x2 = 0.9,
                                 pred_fix_x4 = 1,
                                 nrunif = 3,
                                 nrcovform = 6)

@@

  simres <- gensim(parfun = simfun,
                   seedconst = 7, nrep = 100, ntree = 100,
                   nsteps = 10, stepsize = 1,
                   formula = y~x1+x2+x3+x4+x5+x6,
                   nobs = 500, testnobs = 500L,
                   tree_minsplit = 20, tree_minbucket = 7, tree_mincrit = 0.95,
                   forest_minsplit = 20, forest_minbucket = 7, forest_mincrit = 0, 
                   forest_mtry = 3,
                   censNO = TRUE,
                   gamboost_cvr = FALSE,
                   eval_disttree = TRUE,
                   eval_distforest = TRUE,
                   eval_gamlss = TRUE,
                   eval_bamlss = FALSE,
                   eval_gamboostLSS = FALSE,
                   eval_randomForest = FALSE,
                   eval_cforest = FALSE,
                   nrunif = 3,
                   nrcovform = 6)

@@

\begin{figure}[t!]
\begin{center}
%\begin{subfigure}{0.6\textwidth}
<<sim_crps, fig=TRUE, echo=FALSE>>=
plot_crps(simres)
@
\caption{\label{plot_sim_crps}Results of the simulation study: CRPS} %\code{disttree} vs. \code{distforest} vs. \code{gamlss} vs. \code{randomForest} vs. %\code{bamlss} vs. \code{gamboostLSS}\\
%right: levelplot of the parameter function for $\kappa=5$}
\end{center}
\end{figure}

@@

\subsection{Setting}
The data set used for training and validation consists of observed daily
precipitation sums provided by the national hydrographical service
(\citealt{ehyd}) and numerical weather forecasts from the {U.S.} national
oceanic and atmospheric administration (NOAA).
Both, observations and weather forecasts are available from 1985 trough the
year 2012, focussing on July only in this article.\\
Observations are available on a daily basis, each observation representing the
total precipitation sum fallen within 24~hours observed at 6~UTC at \textbf{95}
stations all over Tyrol and it's close surrounding. The observations are
undergoing a rigorous quality check before made available to the public.  As
covariates the second generation reforecast data set of the global ensemble
forecast system (GEFS) is used (\citealt{hamill2013}).  This data set consists
of an 11-member ensemble based on a fixed version of the numerical model and a
horizontal grid-spacing of about $50~\times~50 km^2$ initialized at 0~UTC from
Dezember 1984 to precent. Each of the 11 ensemble members uses slightly
different perturbed initial conditions trying to predict the situation specific
uncertenty of the atmospheric state.

@@

In particular, 14 basic forecast variables are considered with up to 12 variations of each of these variables being included as covariates. Since each forecast variable consists of an ensemble of individual predictions, the possible variations include the mean, the standard deviation, the minimum and the maximum over all these individual predictions, noted as ensemble mean, ensemble std dev, ensemble minimum and ensemble maximum. All of these values can be calculated on the sums over a whole day (6--30 UTC) or a 6h time window of the day.\\
Another possible option is to consider the mean, minimum or maximum over 24 hours for each individual prediction and then take the ensemble mean or ensemble standard deviation, noted as ensemble mean of mean/minimal/maximal values or ensemble std dev of mean/minimal/maximal values respectively.\\
Table~\ref{covariates} shows all variables together with the number of variations and the provided variations listed in the last column. Altogether 82 covariates are available in this setting.\\
To remove the skewness of ... (the distribution of variables representing precipitation amounts), a power transformation has been used frequently (Box and Cox 1964). In literature, cubic (\citealt{stidd1973}) or square root (\citealt{hutchinson1998b}) transformations have often been suggested but may vary for different climatic zones or temporal aggregation periods.
The response variable \texttt{tp} (total precipitation) is power transformed by a factor of 1.6. This factor is chosen due to experience and knowledge about possible distributions of precipitation measurements. The resulting transformed variable is expected to be normally distributed and left censored at censoring point 0. The same transformation has been chosen for the variable \texttt{cape} (convective available potential energy).

%\newpage
\begin{table}[t!]
%\resizebox{1.4\textwidth}{!}{
\begin{minipage}{\textwidth}
\begin{tabular}{ l  c  l }
\hline
Basic covariates & {\#} & Variations\\
\hline
\textbf{tp}: total precipitation            & 12 variations:  & ensemble mean of sums over 24h, \\
\hspace*{\fill} power transformed (by 1.6) &    & ensemble std dev of sums over 24h, \\ 
\textbf{cape}: convective                  &    & ensemble minimum of sums over 24h, \\
\hspace*{0.4cm} available potential energy &    & ensemble maximum of sums over 24h\\
\hspace*{\fill}  power transformed (by 1.6)&    & \qquad all for 6--30 UTC\\
\hspace*{\fill}                            &    & ensemble mean of sums over 6h\\
\hspace*{\fill}                            &    & \qquad for 6--12, 12--18, 18--24, 24--30 UTC\\
\hspace*{\fill}                            &    & ensemble std dev of sums over 6h\\
\hspace*{\fill}                            &    & \qquad for 6--12, 12--18, 18--24, 24--30 UTC\\
\hline
\textbf{dswrf}: downwards short ware      & 6 variations: & ensemble mean of mean values, \\
\hspace*{\fill} radiation flux (sunshine) &   & ensemble mean of minimal values,\\
\textbf{msl}: mean sea level pressure     &   &  ensemble mean of maximal values,\\
\textbf{pwat}: preciptable water          &   & ensemble std dev of mean values,\\
\textbf{tmax}: 2m maximum                 &   & ensemble std dev of minimal values,\\
\hspace*{1.15cm} temperature              &   & ensemble std dev of maximal values,\\
\textbf{tcolc}: total column-integrated   &   & \qquad all over 6--30 UTC \\
\hspace*{1.05cm} condensate  &   & \\
\textbf{t500}: temperature   &   & \\
\hspace*{1cm} on 500 hPa     &   & \\
\textbf{t700}: temperature   &   & \\
\hspace*{1cm} on 700 hPa     &   & \\
\textbf{t850}: temperature   &   & \\
\hspace{1cm} on 850 hPa      &   & \\
\hline
\textbf{tdiff500850}: temperature         & 3 variations: & ensemble mean of difference in mean,\\
\hspace*{\fill} difference 500 to 850 hPa &   & ensemble minimum of difference in mean,\\
\textbf{tdiff500700}: temperature         &   & ensemble maximum of difference in mean\\
\hspace*{\fill} difference 500 to 700 hPa &   & \qquad all over 6--30 UTC\\
\textbf{tdiff700850}: temperature         &   & \\
\hspace*{\fill} difference 700 to 850 hPa &   & \\
\hline
\textbf{msl{\_}diff}: mean sea level  & 1 variation: & msl{\_}mean{\_}max - msl{\_}mean{\_}min\\
\hspace{1.65cm} difference            &   & \qquad over 6--30 UTC\\
\hline
\end{tabular}
\caption[Table caption text]{Basic covariates together with the number and the type of variations.}
\label{covariates}
\end{minipage} 
%}
\end{table}

>>>>

After explaining the novel methodology in Section~\ref{Methodology} of this paper and presenting results from a simulation study in Section~\ref{Simulation} the focus will then be put on an application in the field of probabilistic forecasting in Section~\ref{ProbabilisticForecasting}. The new method is applied on precipitation data in a mountainous region with a large number of numerical weather prediction quantities and the results are compared to those of other methods commonly used in this research area. In a first setting only one observation station is taken into consideration. The models are trained on 24 years and prbabilistic predictions of the consecutive 4 years are validated. A brief outlook on the results of this setting is given by Figure~\ref{axams_onestation}. For the 24th of July of 2009, 2010, 2011 and 2012 the density functions of the distributions estimated by a forest model are plotted together with the corresponding observations. This illustrates that well fitting predicted expected values are provided by this model as they are close to the observed values.
Beside the expected values the predictions provided by distributional models offer a wide range of information as they specify the whole distribution and therefore enable further methods of inference and analysis.

@@

  save(res, file = "rain_Axams_pred_24to4.rda")
}

@@

## Variable importance
if(file.exists("vimp_crps.rda")){
  #load("~/svn/partykit/pkg/disttree/inst/draft/vimp_crps.rda")
  load("vimp_crps.rda")
} else {

\begin{figure}[t!]
\begin{center}
<<rain_Axams_pred, fig=TRUE, echo=FALSE, width=6.5, height=4.5>>=
plot(x = x, y = y1, type = "l", col = "red", 
     main = paste0("July ", pday, dayending), ylab = "Density", 
     xlab = expression(Total~precipitation~"["~mm^(1/1.6)~"/"~"24h"~"]"),
     #xlab = expression(Total~precipitation~group("[", mm^(1/1.6)~per~day,"]")),
     #{\frac{1}{1.6}}$,
     ylim = c(0,max(y1, y2, y3, y4, pm1, pm2, pm3, pm4) + 0.01),
     xlim = c(-1.5,8))

lines(x = x, y = y2, type = "l", col = "blue")
lines(x = x, y = y3, type = "l", col = "darkgreen")
lines(x = x, y = y4, type = "l", col = "purple")
legend('topright', c("predicted distribution", "point mass at censoring point", "observation"),
       bty = "n", col = "black", lty = c(1, NA, NA), pch = c(NA, 19, 4), cex = 0.8)

@@

# add labels
text(x = -0.8, y = lh1, labels = "2009", col = "red", cex = 0.8)
text(x = -0.8, y = lh2, labels = "2010", col = "blue", cex = 0.8)
text(x = -0.8, y = lh3, labels = "2011", col = "darkgreen", cex = 0.8)
text(x = -0.8, y = lh4, labels = "2012", col = "purple", cex = 0.8)
@
\caption{\label{axams_onestation}Predictions at station Axams for one day in July of 2009, 2010, 2011 and 2012. Observations and predictions are left censored at 0. The corresponding point mass is shown at the censoring point (0).}
\end{center}
\end{figure}

@@

\subsection{Motivation}
As illustrated in Figure~\ref{Devel_parmod} a lot of development has been made in the field of parametric modeling and a wide range of tools can be applied in research areas such as probabilistic forecasting. But as mentioned before, further features such as the ability to capture non-additive effects or interactions are desirable in many applications which demands for more flexible models. Since tree structured models provide these required features a combination of them with parametric models can offer a framework that is able to deal with a wide range of different modeling situations. In this section we are going to introduce the novel methodology which embeds parametric modeling into the idea of tree structured models step by step. We start with a simple distributional fit in Subsection~\ref{distfit}, then build a distributional tree in Subsection~\ref{disttree} and finally a distributional forest in Subsection~\ref{distforest}.

@@

The distribution of the response variable is a left censored normal distribution with censoring point 0. The corresponding distribution parameters are defined by a parameter function $f$ depending on the covariates and an additional factor $\kappa$, also known as the smoothness constraint, which regulates the steepness of certain sections of the parameter function. For low values of $\kappa$ the resulting parameters change rather slowly over varying values of the covariates while high values of $\kappa$ lead to abrupt changes. \\ 
The location parameter $\mu$ depends on $X_1, X_4$ and $\kappa$ and the scale parameter $\sigma$ depends on $\mu, X_1$ and $\kappa$. The other covariates $X_2, X_3, X_5$ and $X_6$ are used as noise variables.\\
The latent variable $Y^{\ast}$ is generated based on the following setting.
\begin{center}
$$ 
Y^{\ast} \sim \mathcal{N} (f(X_1, X_4, \kappa))
$$
with
$$
f(X_1, X_4, \kappa) = \left(\mu(X_1, X_4, \kappa) , \sigma(\mu,X_1,\kappa)\right)
$$ 
and
\begin{align*}
\mu(X_1, X_4, \kappa) = 0.2 
+ \begin{cases}
    7 \cdot \exp{(-(3 \cdot X_1-1.2)^{(2\cdot \kappa)})},
    & \text{if } X_1 < 0.4 \text{ and } X_4 = 0\\
    7 \cdot \exp{(-(3 \cdot X_1-1.2)^{(2\cdot \kappa)})} \cdot \frac{1}{2} + \frac{7}{2},       
    & \text{if } X_1 \geq 0.4 \text{ and } X_4 = 0\\
    10 \cdot \exp{(-(3 \cdot X_1-1.2)^{(2\cdot \kappa)})},
    & \text{if } X_1 < 0.4 \text{ and } X_4 = 1\\
    13 \cdot \exp{(-(3 \cdot X_1-1.2)^{(2\cdot \kappa)})} \cdot \frac{1}{2} + \frac{7}{2},       
    & \text{if } X_1 \geq 0.4 \text{ and } X_4 = 1
\end{cases}
\end{align*}

$$
\sigma(\mu, X_1,\kappa) = 0.5 + \frac{\mu}{5} + 2 \cdot \mathbf{1}(X_1 < -0.4)
$$
%where $\Lambda$ is the cumulative distribution function of the logistic distribution
%$$
%\Lambda(\omega) = \frac{1}{1 + \exp(-\omega)}
%$$
The response variable $Y$ is defined as 
$$
Y = 
\begin{cases}
    Y^{\ast},& \text{if } Y^{\ast} \geq 0\\
    0,       & \text{else}
\end{cases}
$$
\end{center}

To illustrate the parameter functions and the impact of $\kappa$ together with the results of the considered methods they are first applied on one generated data set only where $\kappa$ is set to~1 and then to a second data set where $\kappa$ is set to 10. In Figure~\ref{plot_sim_oneset_mu} the location parameter function $\mu(X_1, X_4, \kappa)$ is plotted along the covariate $X_1$ together with the estimated values of the different models for fixed covariates $x_4 = 0$ and $(x_2, x_3, x_5, x_6) = (0,0,0,0)$. \\

@@

  #define parameter function
  #kappa <- 1
  fun1 <- function(x){
    mu <- 0.2 + 
      (7 * (exp(-(3*x[,1]-1.2)^(2*1)))) * as.numeric(x[,1] < 0.4) + 
      (7 * (exp(-(3*x[,1]-1.2)^(2*1))) * (1/2) + 7/2) * as.numeric(x[,1]>= 0.4) +
      (3 * (exp(-(3*x[,1]-1.2)^(2*1)))) * x[,4]
    sigma <- 0.5 + 
      mu/5 + 
      #(x[,1]>0.4)  #*x[,4]#+
      2*(x[,1] < -0.4) 
      #plogis((1^(1.8)) * 4 * (x[,2] + 0.7)) #* x[,4]
      #4*(1-plogis((1^(1.8)) * 2 * (x[,1] + 0.5))) * x[,4]
      #5 * abs(x[,1]) * x[,4]
    par <- cbind(mu, sigma)
    return(par)
  }

@@

  oneset1 <- sim_oneset_vary_x1(parfun = fun1, nobs = 500, seedconst = 74, ntree = 100,
                                 formula = y~x1+x2+x3+x4+x5+x6, 
                                 tree_minsplit = 20, tree_minbucket = 7, tree_mincrit = 0.95, 
                                 forest_minsplit = 20, forest_minbucket = 7, forest_mincrit = 0, 
                                 forest_mtry = 3, type.tree = "ctree", 
                                 censNO = TRUE, 
                                 pred_fix_x2 = 0.9,
                                 pred_fix_x4 = 0,
                                 nrunif = 3,
                                 nrcovform = 6)

@@

  #define parameter function
  #kappa <- 10
  fun10 <- function(x){
    mu <- 0.2 + 
      (7 * (exp(-(3*x[,1]-1.2)^(2*10)))) * as.numeric(x[,1] < 0.4) + 
      (7 * (exp(-(3*x[,1]-1.2)^(2*10))) * (1/2) + 7/2) * as.numeric(x[,1]>= 0.4) +
      (3 * (exp(-(3*x[,1]-1.2)^(2*10)))) * x[,4]
    sigma <- 0.5 + 
      mu/5 + 
      #(x[,1]>0.4)  #*x[,4]#+
      2*(x[,1] < -0.4) 
      #plogis((10^(1.8)) * 4 * (x[,2] + 0.7)) #* x[,4]
      #4*(1-plogis((10^(1.8)) * 2 * (x[,1] + 0.5))) * x[,4]
      #5 * abs(x[,1]) * x[,4]
    par <- cbind(mu, sigma)
    return(par)
  }

@@

  oneset10 <- sim_oneset_vary_x1(parfun = fun10, nobs = 500, seedconst = 74, ntree = 100,
                                 formula = y~x1+x2+x3+x4+x5+x6, 
                                 tree_minsplit = 20, tree_minbucket = 7, tree_mincrit = 0.95, 
                                 forest_minsplit = 20, forest_minbucket = 7, forest_mincrit = 0, 
                                 forest_mtry = 3, type.tree = "ctree", 
                                 censNO = TRUE, 
                                 pred_fix_x2 = 0.9,
                                 pred_fix_x4 = 0,
                                 nrunif = 3,
                                 nrcovform = 6)

@@

  simres <- gensim(parfun = simfun,
                   seedconst = 74, nrep = 100, ntree = 100,
                   nsteps = 10, stepsize = 1,
                   formula = y~x1+x2+x3+x4+x5+x6,
                   nobs = 500, testnobs = 500L,
                   tree_minsplit = 20, tree_minbucket = 7, tree_mincrit = 0.95,
                   forest_minsplit = 20, forest_minbucket = 7, forest_mincrit = 0, 
                   forest_mtry = 3,
                   censNO = TRUE,
                   gamboost_cvr = FALSE,
                   eval_disttree = TRUE,
                   eval_distforest = TRUE,
                   eval_gamlss = TRUE,
                   eval_bamlss = FALSE,
                   eval_gamboostLSS = FALSE,
                   eval_randomForest = FALSE,
                   eval_cforest = FALSE,
                   nrunif = 3,
                   nrcovform = 6)

@@

\begin{figure}[t!]
\begin{center}
%\begin{subfigure}{0.6\textwidth}
<<sim_crps, fig=TRUE, echo=FALSE>>=
plot_crps(simres)
@
\caption{\label{plot_sim_crps}CRPS of the three models evaluated on various data sets over increasing $\kappa$} %\code{disttree} vs. \code{distforest} vs. \code{gamlss} vs. \code{randomForest} vs. %\code{bamlss} vs. \code{gamboostLSS}\\
%right: levelplot of the parameter function for $\kappa=5$}
\end{center}
\end{figure}

@@

\subsection{Setting}
The data set used for training and validation consists of observed daily precipitation sums provided by the National Hydrographical Service (\citealt{ehyd}) and numerical weather forecasts from the {U.S.} National Oceanic and Atmospheric Administration (NOAA).
Both, observations and weather forecasts are available from 1985 trough the year 2012, focussing on July only in this article.\\ 
Observations are available on a daily basis, each observation representing the total precipitation sum fallen within 24~hours observed at 6~UTC at 95 stations all over Tyrol and it's close surrounding. The observations are undergoing a rigorous quality check before made available to the public.  As covariates the second generation reforecast data set of the global ensemble forecast system (GEFS) is used (\citealt{hamill2013}).  This data set consists of an 11-member ensemble based on a fixed version of the numerical model and a horizontal grid-spacing of about $50~\times~50 \text{km}^2$ initialized at 0~UTC from Dezember 1984 to precent. Each of the 11 ensemble members uses slightly different perturbed initial conditions trying to predict the situation specific uncertenty of the atmospheric state.

@@

In particular, 14 basic forecast variables are considered with up to 12 variations of each of these variables being included as covariates. Since each forecast variable consists of an ensemble of individual predictions, the possible variations include the mean, the standard deviation, the minimum and the maximum over all ensemble members, noted as ensemble mean, ensemble std dev, ensemble minimum and ensemble maximum. All of these values can be calculated on the sums over 24 hours (6--30 UTC) or a 6 hour time window.\\
Another possible option is to consider the mean, minimum or maximum over 24 hours for each individual prediction and then take the ensemble mean or ensemble standard deviation, noted as ensemble mean of mean/minimal/maximal values or ensemble std dev of mean/minimal/maximal values respectively.\\
Table~\ref{covariates} shows all variables together with the number of variations and the provided variations listed in the last column. Altogether 82 covariates are available in this setting.\\
To remove large parts of the skewness of precipitation data, a power transformation has been used frequently (Box and Cox 1964). In literature, cubic (\citealt{stidd1973}) or square root (\citealt{hutchinson1998b}) transformations have often been suggested but may vary for different climatic zones or temporal aggregation periods. In \citealt{stauffer2017a} optimization methods have been applied to find the best transformation factor which resulted in a value of $\frac{1}{1.6}$. Since the data used for this optimization covers almost the same geographical region as considered in this article the same factor is chosen here for the power transformation of the response variable \texttt{tp} (total precipitation). The resulting transformed variable is expected to be normally distributed and left censored at censoring point 0. For simplicity the same transformation has been chosen for the variable \texttt{cape} (convective available potential energy).

%\newpage
\begin{table}[t!]
%\resizebox{1.4\textwidth}{!}{
\begin{minipage}{\textwidth}
\begin{tabular}{ l  c  l }
\hline
Basic covariates & {\#} & Variations\\
\hline
\textbf{tp}: total precipitation            & 12 variations:  & ensemble mean of sums over 24h, \\
\hspace*{\fill} power transformed (by $\frac{1}{1.6}$) &    & ensemble std dev of sums over 24h, \\ 
\textbf{cape}: convective                  &    & ensemble minimum of sums over 24h, \\
\hspace*{0.4cm} available potential energy &    & ensemble maximum of sums over 24h\\
\hspace*{\fill}  power transformed (by $\frac{1}{1.6}$)&    & \qquad all for 6--30 UTC\\
\hspace*{\fill}                            &    & ensemble mean of sums over 6h\\
\hspace*{\fill}                            &    & \qquad for 6--12, 12--18, 18--24, 24--30 UTC\\
\hspace*{\fill}                            &    & ensemble std dev of sums over 6h\\
\hspace*{\fill}                            &    & \qquad for 6--12, 12--18, 18--24, 24--30 UTC\\
\hline
\textbf{dswrf}: downwards short ware      & 6 variations: & ensemble mean of mean values, \\
\hspace*{\fill} radiation flux (``sunshine'') &   & ensemble mean of minimal values,\\
\textbf{msl}: mean sea level pressure     &   &  ensemble mean of maximal values,\\
\textbf{pwat}: preciptable water          &   & ensemble std dev of mean values,\\
\textbf{tmax}: 2m maximum                 &   & ensemble std dev of minimal values,\\
\hspace*{1.15cm} temperature              &   & ensemble std dev of maximal values,\\
\textbf{tcolc}: total column-integrated   &   & \qquad all over 6--30 UTC \\
\hspace*{1.05cm} condensate  &   & \\
\textbf{t500}: temperature   &   & \\
\hspace*{1cm} on 500 hPa     &   & \\
\textbf{t700}: temperature   &   & \\
\hspace*{1cm} on 700 hPa     &   & \\
\textbf{t850}: temperature   &   & \\
\hspace{1cm} on 850 hPa      &   & \\
\hline
\textbf{tdiff500850}: temperature         & 3 variations: & ensemble mean of difference in mean,\\
\hspace*{\fill} difference 500 to 850 hPa &   & ensemble minimum of difference in mean,\\
\textbf{tdiff500700}: temperature         &   & ensemble maximum of difference in mean\\
\hspace*{\fill} difference 500 to 700 hPa &   & \qquad all over 6--30 UTC\\
\textbf{tdiff700850}: temperature         &   & \\
\hspace*{\fill} difference 700 to 850 hPa &   & \\
\hline
\textbf{msl{\_}diff}: mean sea level  & 1 variation: & msl{\_}mean{\_}max - msl{\_}mean{\_}min\\
\hspace{1.65cm} difference            &   & \qquad over 6--30 UTC\\
\hline
\end{tabular}
\caption[Table caption text]{Basic covariates together with the number and the type of variations.}
\label{covariates}
\end{minipage} 
%}
\end{table}

\subsection{Models}
Next to distributional forests three other methods are applied to this data set and their results are compared based on CRPS values. One of the reference models is an EMOS models as this is a common choice in this type of setting / research field. The other two models are both GAMs, once in a prespecified form in terms of variable selection for each distribution parameters and once applying a boosting algorithm). 
The chosen methods and their specifications are listed in Table~\ref{model_specification}.
\begin{table}[t!]
%\resizebox{1.4\textwidth}{!}{
%\begin{minipage}{\textwidth}
\begin{center}
\begin{tabular}{ l  l   l   l }
\hline
 & \textbf{Type} & \textbf{Location ($\boldsymbol{\mu}$)} & \textbf{Scale ($\boldsymbol{\log(\sigma)}$)} \\
\hline
\textbf{Forest} & recursive    & all & all \\ 
                & partitioning &     &     \\ 
\hline
\textbf{prespecified GAM} & spline  & tp{\_}mean, & tppow{\_}sprd,\\
                            & in each & tp{\_}max,  & dswrf{\_}sprd{\_}mean,\\
 & & tp{\_}mean1218 $\ast$ & tp{\_}sprd1218 $\ast$\\ 
 & & \quad cape{\_}mean1218, & \quad  cape{\_}mean1218,\\
 & & dswrf{\_}mean{\_}mean, & tcolc{\_}sprd{\_}mean,\\
 & & tcolc{\_}mean{\_}mean, & tdiff500850{\_}mean\\
 & & pwat{\_}mean{\_}mean, & \\
 & & tdiff500850{\_}mean, & \\
 & & msl{\_}diff & \\
\hline
\textbf{boosted GAM} & spline  & all & all \\
                       & in each &     &     \\
\hline
\textbf{EMOS} & linear & tp{\_}mean & tp{\_}spread \\
\hline
\end{tabular}
\caption[Table caption text]{Table of applied methods together with their type (way of including the covariates) and the included covariates for each distribution parameter.}
\label{model_specification}
%}
\end{center}
\end{table}
